Improving Dynamic Kidney Exchange

### 2018-04-06 chapter2/chapter2.tex
82 pages

%It includes the rules for which pairs are deemed compatible, the entry and death processes that govern the evolution of the pool, which characteristics are observable to the algorithm at each period, and the reward observed by the agent as sequentially matches and clears pairs. More rigorously, at each period an environment is described bu tuple $(G, u)$, where:

% 
% \begin{itemize}
%   \item $G = (V, A, X, \pi)$ is a graph with vertex set $V$, adjacency matrix $A$, vertex characteristic matrix $X$, and $\pi: \mathbb{N} \times V \rightarrow \{0, 1\}$ is a function that maps a period $t \in \mathbb{N}$ and a vertex $v \in V$ to whether that vertex was alive on period $t$ (1) or not (0). \footnote{We decided to model dynamic graphs in this manner as it is closer to our simulation code. Note that all the information about entries and exits can be expressed by the function $\pi$.}
%   \item $u: 2^V \rightarrow \mathbb{R} \cup \{-\infty\} $ is the reward associated with matching subset of $V$. Infeasible exchanges receive reward $-\infty$.
% \end{itemize}

@@

\begin{itemize} 
  \item Patients that were registered for more than one organ (including those who were simultaneously waiting for kidney and pancreas)
  \item Patients that were not waiting for their first kidney transplant
  \item Donors and patients with incomplete tissue-type profile information.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
\caption{\textbf{Patient and donor cPRA} \\
    Computed from historical data. Our patient cPRA is the frequency of living donors that exhibit antigens that are unacceptable to the patient. We also define a donor cPRA by calculating the frequency of patients that have antibodies against the donors antigens.}
\label{fig:cpra}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\hspace*{-3.5cm}
\includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
\caption{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
Ratio of average per-period matched pairs for different entry and death rates, over 3000 periods (darker hues are better). \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
\label{fig:greedy_opt_comparison}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
\caption{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles. See Algorithm \ref{alg:direct_prediction} for details.}
\label{fig:direct_prediction}
\end{figure}

@@

 \begin{figure}[htbp]
  \centering
  %\hspace{-1cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
  \caption{\textbf{Multi-armed bandit methods} \\ 
  One period of simulation using multi-armed bandit methods to choose cycles.}
  \label{fig:mab}
  \end{figure}

  \begin{figure}[htbp]
  \centering
  %\hspace*{-0.5cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
  \caption{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future. Such cycles get a \emph{lower} reward, and are left for later.}
  \label{fig:get_pseudo_reward}
  \end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_grb}
\caption{\textbf{Performance of gradient boosting in direct prediction method} \\
    Ratio of average matched patients against \textsc{Myopic}. Simulations ran over 750 periods (after a 250-period burn-in sequence). Lower row is when data was augmented with information about graph.}
\label{fig:traditional_ml_mdp_algo_grb}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_lr}
\caption{\textbf{Performance of logistic regression in direct prediction method} \\
    Similar to gradient boosting shown in Figure \ref{fig:traditional_ml_mdp_algo_grb}, logistic regression is only able to perform better than \textsc{Myopic} when the death rate is large.}
\label{fig:traditional_ml_mdp_algo_lr}
\end{figure}
bb
\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_rf}
\caption{\textbf{Performance of random forests in direct prediction method} \\
    Random Forests's poor performance is likely due to its low accuracy, as we show on Table \ref{tab:tradional_ml_classifier}.}
\label{fig:traditional_ml_mdp_algo_rf}
\end{figure}

@@

\begin{figure}
\centering
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_ABO.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_RSU.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_OPTN.pdf}
\caption{\textbf{Performance multi-armed bandit methods across environments} \\
Average percent improvement by the \emph{multi-armed bandit} method over \textsc{Myopic}. Averages were taken over 750 periods, after a 250-period burn-in sequence. See also Tables \ref{tab:ucb1}-\ref{tab:exp3}. \\
{\scriptsize Note: In $RSU$ row, white entries are missing (they are not zero).}}
\label{fig:mab}
\end{figure}

@@

@@

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_ml_classifier.tex}
   \caption{\textbf{Performance of statistical methods as classifiers} \\ 
   We experiment with three variations on the \emph{direct prediction} method by implementing it using different statistical. Here we see their performance as classifiers, that is, how well they are able to predict that the node was chosen to be matched by an offline algorithm. \\
  }
  \label{tab:traditional_ml_classifier}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_none.tex}
   \caption{\textbf{Gradient Boosting (no graph information)} \\ }
  \label{tab:grb_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_networkx.tex}
   \caption{\textbf{Gradient Boosting (augmented with graph information)} \\ }
  \label{tab:grb_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_none.tex}
   \caption{\textbf{Logistic Regression (no graph information)} \\ }
  \label{tab:lr_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_networkx.tex}
   \caption{\textbf{Logistic Regression (augmented with graph information)} \\ }
  \label{tab:lr_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_none.tex}
   \caption{\textbf{Random Forests (no graph information)} \\ }
  \label{tab:rf_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_networkx.tex}
   \caption{\textbf{Random Forests (augmented with graph information)} \\ }
  \label{tab:rf_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_UCB1.tex}
   \caption{\textbf{Multi-armed bandit algorithm UCB1} \\ 
  \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration. Tables for other bandits are similar, and can be found in the online supplement that will be available online.}
  \label{tab:ucb1}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_Thompson.tex}
  \caption{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
 \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration.}
  \label{tab:thompson}
\end{table}

>>>>

@@

\begin{itemize} 
  \item Patients that were registered for more than one organ (including those who were simultaneously waiting for kidney and pancreas)
  \item Patients that were not waiting for their first kidney transplant
  \item Donors and patients with incomplete tissue-type profile information.
\end{itemize}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/computed_cPRA.pdf}
\caption[Patient and donor cPRA]{\textbf{Patient and donor cPRA} \\
    Computed from historical data. Our patient cPRA is the frequency of living donors that exhibit antigens that are unacceptable to the patient. We also define a donor cPRA by calculating the frequency of patients that have antibodies against the donors antigens.}
\label{fig:cpra}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\hspace*{-3.5cm}
\includegraphics[width=1.4\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/greedy_opt_comparison_mcl2_2.pdf}
\caption[Comparing \textsc{Myopic} and \textsc{OPT}]{\textbf{Comparing \textsc{Myopic} and \textsc{OPT}}  \\
Ratio of average per-period matched pairs for different entry and death rates, over 3000 periods (darker hues are better). \texttt{Myopic} has better chances of achieving performances similar to \texttt{OPT} when the death rate is high (moving rightwards on the graphs), or when entry rate is high (moving downwards on the graphs).}
\label{fig:greedy_opt_comparison}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/direct_prediction.pdf}.
\caption[Direct prediction method]{\textbf{Direct prediction method} \\ One period of simulation using direct prediction methods to choose cycles. See Algorithm \ref{alg:direct_prediction} for details.}
\label{fig:direct_prediction}
\end{figure}

@@

 \begin{figure}[htbp]
  \centering
  %\hspace{-1cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab.pdf}
  \caption[Multi-armed bandit methods (one-period)]{\textbf{Multi-armed bandit methods} \\ 
  One period of simulation using multi-armed bandit methods to choose cycles.}
  \label{fig:mab}
  \end{figure}

  \begin{figure}[htbp]
  \centering
  %\hspace*{-0.5cm}
  \includegraphics[width=1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_step.pdf}
  \caption[Function \textsc{get\_pseudo\_reward}]{\textbf{Function} \textsc{get\_pseudo\_reward} \\ Multi-armed bandits evaluate if a cycle should be cleared by checking if there is a high chance that the cycle will be used in the future. Such cycles get a \emph{lower} reward, and are left for later.}
  \label{fig:get_pseudo_reward}
  \end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_grb}
\caption[Performance of gradient boosting in direct prediction method]{\textbf{Performance of gradient boosting in direct prediction method} \\
    Ratio of average matched patients against \textsc{Myopic}. Simulations ran over 750 periods (after a 250-period burn-in sequence). Lower row is when data was augmented with information about graph.}
\label{fig:traditional_ml_mdp_algo_grb}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_lr}
\caption[Performance of logistic regression in direct prediction method]{\textbf{Performance of logistic regression in direct prediction method} \\
    Similar to gradient boosting shown in Figure \ref{fig:traditional_ml_mdp_algo_grb}, logistic regression is only able to perform better than \textsc{Myopic} when the death rate is large.}
\label{fig:traditional_ml_mdp_algo_lr}
\end{figure}
bb
\begin{figure}
\centering
\includegraphics[width=\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/traditional_ml_mdp_algo_rf}
\caption[Performance of random forests in direct prediction method]{\textbf{Performance of random forests in direct prediction method} \\
    Random Forests's poor performance is likely due to its low accuracy, as we show on Table \ref{tab:tradional_ml_classifier}.}
\label{fig:traditional_ml_mdp_algo_rf}
\end{figure}

@@

\begin{figure}
\centering
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_ABO.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_RSU.pdf}
\hspace*{-1.6cm}
\includegraphics[width=1.1\textwidth]{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/figures/mab_OPTN.pdf}
\caption[Performance of multi-armed bandit across environments]{\textbf{Performance multi-armed bandit methods across environments} \\
Average percent improvement by the \emph{multi-armed bandit} method over \textsc{Myopic}. Averages were taken over 750 periods, after a 250-period burn-in sequence. See also Tables \ref{tab:ucb1}-\ref{tab:exp3}. \\
{\scriptsize Note: In $RSU$ row, white entries are missing (they are not zero).}}
\label{fig:mab}
\end{figure}

@@

[[[Other loss functions]]]

@@

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_ml_classifier.tex}
   \caption[Performance of statistical methods as classifiers]{
   \textbf{Performance of statistical methods as classifiers} \\ 
   We experiment with three variations on the \emph{direct prediction} method by implementing it using different statistical. Here we see their performance as classifiers, that is, how well they are able to predict that the node was chosen to be matched by an offline algorithm. \\
  }
  \label{tab:traditional_ml_classifier}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_none.tex}
  \caption[Performance of direct prediction method with gradient boosting (1)]{\textbf{Performance of direct prediction method with gradient boosting} Not using information about graph.}
  \label{tab:grb_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_grb_networkx.tex}
  \caption[Performance of direct prediction method with gradient boosting (2)]{\textbf{Performance of direct prediction method with gradient boosting} Using information about graph.}
  \label{tab:grb_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_none.tex}
  \caption[Performance of direct prediction method with logistic regression (1)]{\textbf{Performance of direct prediction method with logistic regression} Not using information about graph.}
  \label{tab:lr_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_lr_networkx.tex}
  \caption[Performance of direct prediction method with logistic regression (2)]{\textbf{Performance of direct prediction method with logistic regression} Using information about graph.}
  \label{tab:lr_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_none.tex}
  \caption[Performance of direct prediction method with random forests (1)]{\textbf{Performance of direct prediction method with random forests} Not using information about graph.}
  \label{tab:rf_none}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/traditional_mdp_rf_networkx.tex}
  \caption[Performance of direct prediction method with random forests (2)]{\textbf{Performance of direct prediction method with random forests} Using information about graph.}
  \label{tab:rf_networkx}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_UCB1.tex}
   \caption[Performance of multi-armed bandit method using UCB1]{\textbf{Multi-armed bandit algorithm UCB1} \\ 
  \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration. Tables for other bandits are similar, and can be found in the online supplement that will be available online.}
  \label{tab:ucb1}
\end{table}

\begin{table}
  \centering
  \singlespacing
  \hspace*{-0.5cm}
  \input{/Users/vitorhadad/Documents/kidney/matching/phd_thesis/chapter2/tables/mab_Thompson.tex}
   \caption[Performance of multi-armed bandit method using Thompson sampling]{\textbf{Multi-armed bandit algorithm Thompson Sampling} \\ 
 \textsc{Environ} is the environment used for simulation. \textsc{Entry} and \textsc{Death} are the Poisson entry rate of entry and the the Geometric rate of departure (times 100), respectively.  \textsc{Mean} and \textsc{Std} refers to the average number of matched patients over 750 periods (after 250 periods of burn-in). \textsc{Difference} and \textsc{Ratio} compare the average improvement between the algorithm and \textsc{Myopic}. \textsc{N} is the number of 1000-period simulations that used that particular configuration.}
  \label{tab:thompson}
\end{table}
